Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-09-27 paper_comparison/paper_comparison/paper_comparison.Rmd
modify the example table

Residuals might be plotted, as a histogram or quantile-quantile plot to examine the distribution. Using the classical linear regression model as an example, if the distribution is symmetric and unimodal, it is well-behaved. But if the distribution is skewed, bimodal, multimodal, or contains outliers, there is cause for concern. The distribution could also be inspected by conducting a goodness of fit test, such as the Shapiro-Wilk Normality test [@shapiro1965analysis].

@@

Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals. As shown in Figure \ref{fig:commondepartures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns. There are various types of departures from a good residual plot. We will only discuss three most commonly checked departures in this paper. Namely, non-linearity, heterskedasticity and non-normality.

Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation. Any non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity. At the top-right of Figure [ref here], there is a residual plot giving an example of the visual pattern of non-linearity when a cubic term is not captured by the regression model.

@@

Non-normality is usually harder to be detected from a residual plot compared to non-linearity and heterskedasticity. Besides, a quantile-quantile plot can often do a better job for this task. Note that not all regression models assume normality of the error term, but a certain amount of them does. If this assumption happens to be false, then it is expected to see data points do not distribute noramlly on the y-axis. For example, given a skewed error distribution, one will see fewer data points and more outliers on one side of the zeroline as shown in the bottom-right of Figure \ref{fig:commondepartures}.

@@

data.frame(plot = c("B", "C", "D"), 
           model = c("Non-linearity", "Heteroskedasticity", "Non-normality"),
           f = c(POLY_MODEL$test(dat1)$p_value,
                 POLY_MODEL$test(mutate(dat2, z = poly_model()$gen(300, computed = select(dat2, x))$z))$p_value,
                 POLY_MODEL$test(mutate(dat3, z = poly_model()$gen(300, computed = select(dat3, x))$z))$p_value),
           b = c(HETER_MODEL$test(dat1)$p_value,
                 HETER_MODEL$test(dat2)$p_value,
                 HETER_MODEL$test(dat3)$p_value),
           s = c(shapiro.test(dat1$.resid)$p.value,
                 shapiro.test(dat2$.resid)$p.value,
                 shapiro.test(dat3$.resid)$p.value)) %>%
  mutate(f = format(round(f, digits = 3), nsmall = 3), b = format(round(b, digits = 3), nsmall = 3), s = format(round(s, digits = 3), nsmall = 3)) -> table_dat

@@

  table_dat %>%
  knitr::kable(col.names = c("Residual plots", "Departures", "Model structural", "Breusch-Pagan", "Shapiro–Wilk"),
               align = "llrrr",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Example residual plots examined by the corresponding conventional tests. P-values smaller than 0.05 are showed in bold.', 
               label = "example-residual-plot-table" )

@@

The effectiveness of lineup protocol has already been validated by @majumder_validation_2013 under relatively simple classical normal linear regression model settings with only one or two regressors. Their results suggest visual test is capable of testing the significance of a single regressor with a similar power as a t-test, though they expressed that in general it is unnecessary to use visual inference if there exists a conventional test and they didn't expect the visual test to perform equally well as the conventional test. In their third experiment, where there does not exist a proper conventional test, visual test outperforms the conventional test for a large margin. This is encouraging as it promotes the use of visual inference in border field of data science where there are no existing statistical testing procedures. In fact, lineup protocol has been integrated into some model diagnostic tools such as @loy2013diagnostic.

@@

```{r residual-plot-cubic-heter, fig.cap='Residuals vs. fitted values plot for a classical linear regression model. The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model. From the residual plot, "butterfly shape" can be observed which generally would be interpretd as evidence of heteroskedasticity. Further, from the outline of the shape, nonlinear patterns exist. Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process. \\label{fig:residual-plot-cubic-heter}', eval = FALSE}

@@

```{r example-lineup, fig.cap="Lineup 24 in experiment I. Can you spot the most different plot? \\label{fig:example-lineup}"}
VI_MODEL$plot_lineup(polynomials_lineup[[24]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

>>>>

Residuals might be plotted, as a histogram or quantile-quantile plot to examine the distribution. Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, it is well-behaved. But if the distribution is skewed, bimodal, multimodal, or contains outliers, there is cause for concern. The distribution could also be inspected by conducting a goodness of fit test, such as the Shapiro-Wilk Normality test [@shapiro1965analysis].

@@

Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals. As shown in Figure \ref{fig:commondepartures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns. There are various types of departures from a good residual plot. Two most commonly checked departures - non-linearity, heterskedasticity and non-normality will be discussed in this paper.

Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation. Any non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity. An example residual plot containing visual pattern of non-linearity is given at the top-right of Figure [ref here]. One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.

@@

Compared to non-linearity and heterskedasticity, non-normality is usually harder to be detected from a residual plot since scatter plot is not excel in revealing marginal distribution. A favourable graphical summarise for this task is the quantile-quantile plot. However, for a consistent comparison in the later part of this paper, residual plot will be the focus. Besides, it is important to note that not all regression models assume normality for the error term, but a certain amount do including the classical normal linear regression model. In the case that the normality assumption doesn't hold, it is expected to observe data points that do not centralize around the horizontal zero line and unevenly distribute at both sides of the zero line. For example, given a skewed error distribution, fewer data points and more outliers are on one side of the zero line as shown at the bottom-right of Figure \ref{fig:commondepartures}.

@@

data.frame(plot = c("A", "B", "C", "D"), 
           model = c("None", "Non-linearity", "Heteroskedasticity", "Non-normality"),
           f = c(POLY_MODEL$test(dat)$p_value, 
                 POLY_MODEL$test(dat1)$p_value,
                 POLY_MODEL$test(mutate(dat2, z = poly_model()$gen(300, computed = select(dat2, x))$z))$p_value,
                 POLY_MODEL$test(mutate(dat3, z = poly_model()$gen(300, computed = select(dat3, x))$z))$p_value),
           b = c(HETER_MODEL$test(dat)$p_value, 
                 HETER_MODEL$test(dat1)$p_value,
                 HETER_MODEL$test(dat2)$p_value,
                 HETER_MODEL$test(dat3)$p_value),
           s = c(shapiro.test(dat$.resid)$p.value, 
                 shapiro.test(dat1$.resid)$p.value,
                 shapiro.test(dat2$.resid)$p.value,
                 shapiro.test(dat3$.resid)$p.value)) %>%
  mutate(f = format(round(f, digits = 3), nsmall = 3), b = format(round(b, digits = 3), nsmall = 3), s = format(round(s, digits = 3), nsmall = 3)) -> table_dat

@@

  table_dat %>%
  knitr::kable(col.names = c("Plot", "Departures", "Model structural", "Breusch-Pagan", "Shapiro–Wilk"),
               align = "llrrr",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Results of example residual plots given in Figure \\ref{fig:commondepartures} examined by the Model structural test (F-test), the Breusch-Pagan test and the Shapiro–Wilk test. P-values smaller than 0.05 are showed in bold.', 
               label = "example-residual-plot-table" )

@@

The effectiveness of lineup protocol has already been validated by @majumder_validation_2013 under relatively simple classical normal linear regression model settings with only one or two regressors. Their results suggest visual test is capable of testing the significance of a single regressor with a similar power as a t-test, though they expressed that in general it is unnecessary to use visual inference if there exists a conventional test and they didn't expect the visual test to perform equally well as the conventional test. In their third experiment, where there does not exist a proper conventional test, visual test outperforms the conventional test for a large margin. This is encouraging as it promotes the use of visual inference in border field of data science where there are no existing statistical testing procedures. In fact, lineup protocol has been integrated successfully into diagnostic tools of hierarchical linear models [@loy2013diagnostic].

@@

```{r residual-plot-cubic-heter, fig.cap='Residuals vs. fitted values plot for a classical normal linear regression model. The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model. From the residual plot, "butterfly shape" can be observed which generally would be interpretd as evidence of heteroskedasticity. Further, from the outline of the shape, nonlinear patterns exist. Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process. \\label{fig:residual-plot-cubic-heter}', eval = FALSE}

@@

```{r example-lineup, fig.cap="Lineup 24 in experiment I. Can you spot the most different plot? \\label{fig:example-lineup}"}
VI_MODEL$plot_lineup(vi_lineup[["poly_24"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

### 2022-09-27 paper_comparison/paper_comparison/paper_comparison.tex
modify the example table

Residuals might be plotted, as a histogram or quantile-quantile plot to
examine the distribution. Using the classical linear regression model as
an example, if the distribution is symmetric and unimodal, it is
well-behaved. But if the distribution is skewed, bimodal, multimodal, or
contains outliers, there is cause for concern. The distribution could
also be inspected by conducting a goodness of fit test, such as the
Shapiro-Wilk Normality test \citep{shapiro1965analysis}.

@@

Graphical summaries in which residuals are plotted against fitted values
or other functions of the predictor variables that are approximately
orthogonal to residuals are referred to as standard residual plots in
\citet{cook1982residuals}. As shown in Figure
\ref{fig:commondepartures}, the top-left panel is a good residual plot
with residuals evenly distributed at both sides of the horizontal zero
line showing no noticeable patterns. There are various types of
departures from a good residual plot. We will only discuss three most
commonly checked departures in this paper. Namely, non-linearity,
heterskedasticity and non-normality.

Non-linearity is a type of model misspecification caused by failing to
include higher order terms of the regressors in the regression equation.
Any non-linear functional form of residuals on fitted values presented
in the residual plot could be considered as an indicative of
non-linearity. At the top-right of Figure {[}ref here{]}, there is a
residual plot giving an example of the visual pattern of non-linearity
when a cubic term is not captured by the regression model.

@@

\caption{\label{tab:example-residual-plot-table}Example residual plots examined by the corresponding conventional tests. P-values smaller than 0.05 are showed in bold.}
\centering
\begin{tabular}[t]{llrrr}
\toprule
Residual plots & Departures & Model structural & Breusch-Pagan & Shapiro–Wilk\\
\midrule
B & Non-linearity & \textbf{0.000} & \textbf{0.000} & \textbf{0.039}\\
C & Heteroskedasticity & 0.378 & \textbf{0.000} & \textbf{0.000}\\
D & Non-normality & 0.667 & 0.736 & \textbf{0.000}\\
\bottomrule
\end{tabular}
\end{table}

@@

The effectiveness of lineup protocol has already been validated by
\citet{majumder_validation_2013} under relatively simple classical
normal linear regression model settings with only one or two regressors.
Their results suggest visual test is capable of testing the significance
of a single regressor with a similar power as a t-test, though they
expressed that in general it is unnecessary to use visual inference if
there exists a conventional test and they didn't expect the visual test
to perform equally well as the conventional test. In their third
experiment, where there does not exist a proper conventional test,
visual test outperforms the conventional test for a large margin. This
is encouraging as it promotes the use of visual inference in border
field of data science where there are no existing statistical testing
procedures. In fact, lineup protocol has been integrated into some model
diagnostic tools such as \citet{loy2013diagnostic}.

>>>>

Residuals might be plotted, as a histogram or quantile-quantile plot to
examine the distribution. Using the classical normal linear regression
model as an example, if the distribution is symmetric and unimodal, it
is well-behaved. But if the distribution is skewed, bimodal, multimodal,
or contains outliers, there is cause for concern. The distribution could
also be inspected by conducting a goodness of fit test, such as the
Shapiro-Wilk Normality test \citep{shapiro1965analysis}.

@@

Graphical summaries in which residuals are plotted against fitted values
or other functions of the predictor variables that are approximately
orthogonal to residuals are referred to as standard residual plots in
\citet{cook1982residuals}. As shown in Figure
\ref{fig:commondepartures}, the top-left panel is a good residual plot
with residuals evenly distributed at both sides of the horizontal zero
line showing no noticeable patterns. There are various types of
departures from a good residual plot. Two most commonly checked
departures - non-linearity, heterskedasticity and non-normality will be
discussed in this paper.

Non-linearity is a type of model misspecification caused by failing to
include higher order terms of the regressors in the regression equation.
Any non-linear functional form of residuals on fitted values presented
in the residual plot could be considered as an indicative of
non-linearity. An example residual plot containing visual pattern of
non-linearity is given at the top-right of Figure {[}ref here{]}. One
can clearly observe the ``S-shape'' from the residual plot as the cubic
term is not captured by the misspecified model.

@@

\caption{\label{tab:example-residual-plot-table}Results of example residual plots given in Figure \ref{fig:commondepartures} examined by the Model structural test (F-test), the Breusch-Pagan test and the Shapiro–Wilk test. P-values smaller than 0.05 are showed in bold.}
\centering
\begin{tabular}[t]{llrrr}
\toprule
Plot & Departures & Model structural & Breusch-Pagan & Shapiro–Wilk\\
\midrule
A & None & 0.434 & 0.133 & 0.728\\
B & Non-linearity & \textbf{0.000} & \textbf{0.000} & \textbf{0.039}\\
C & Heteroskedasticity & 0.378 & \textbf{0.000} & \textbf{0.000}\\
D & Non-normality & 0.667 & 0.736 & \textbf{0.000}\\
\bottomrule
\end{tabular}
\end{table}

@@

The effectiveness of lineup protocol has already been validated by
\citet{majumder_validation_2013} under relatively simple classical
normal linear regression model settings with only one or two regressors.
Their results suggest visual test is capable of testing the significance
of a single regressor with a similar power as a t-test, though they
expressed that in general it is unnecessary to use visual inference if
there exists a conventional test and they didn't expect the visual test
to perform equally well as the conventional test. In their third
experiment, where there does not exist a proper conventional test,
visual test outperforms the conventional test for a large margin. This
is encouraging as it promotes the use of visual inference in border
field of data science where there are no existing statistical testing
procedures. In fact, lineup protocol has been integrated successfully
into diagnostic tools of hierarchical linear models
\citep{loy2013diagnostic}.
